---
title: "Fandango_Notebook"
author: "Neil Yetz & Gemma Wallace"
subtitle: PSY 652 RM Module 5 Lab
output:
  html_notebook:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
---
This activity is designed to build familiarity with data wrangling techniques in R.
In this activity you will create a new R notebook using the Fandango dataset. This is the raw data behind the story "Be Suspicious Of Online Movie Ratings, Especially Fandango's. (http://fivethirtyeight.com/features/fandango-movies-ratings/).”  The dataframe contains every film that has a Rotten Tomatoes rating, a RT User rating, a Metacritic score, a Metacritic User score, and IMDb score, and at least 30 fan reviews on Fandango.  

# Load libraries
```{r, message = FALSE}
library(tidyverse)
```

# Import data
```{r, message = FALSE}
fandango <- read_csv("fandango_module5lab.csv")
```

# See the structure of the data frame
```{r}
str(fandango)
```

# Examine descriptives using the summary function
```{r}
summary(fandango)
```

# A collection of data wrangling techniques

## filter the data
The filter function allows you subset data by rows.
```{r}
film_2014 <- filter(fandango, year==2014)
film_2014 <- filter(fandango, year==2014 & stars >=4)
```

## select the data
The select function allows you to subset data by column.
```{r}
fandango_subset <- select(fandango, year, stars)
```

## mutate the data
The mutate function allows you to create new variables in your dataset. The mutate function can either create new variables or modify existing variables.
```{r}
fandango <- mutate(fandango, stars_ten = stars*10)
```

## arrange the data
The arrange function allows you to sort the data by certain variable(s).
```{r}
fandango_sorted <- arrange(fandango, year)
```

## summarize the data
The summarize function allows you to collapse cases for easier descriptive interpretations and plotting.
```{r}
fandango_year <- group_by(fandango,year)
summarize(fandango_year, year_mean = mean(stars))
```
