---
title: "R Notebook"
output: html_notebook
---

# Load libraries
```{r, message = FALSE}
library(tidyverse)
library(psych)
```

# Import data
```{r,message=FALSE}
slp <- read_csv("slpdata.csv")
```

# Reduce to variables of interest
```{r}
slp_reduced <- select(slp, cond, sleep)
```


# Factor the grouping variable
```{r}
slp_reduced <- slp_reduced %>% 
  mutate(cond.f = factor(cond, 
                                levels = c(1,2,3),
                                labels = c("Control", "Sleep int", "Sleep int + Partner")))
```


# Get descriptives
## Overall
```{r}

describe(slp_reduced)

```

#Grouped
```{r}
slp_grouped <- group_by(slp_reduced, cond.f)
summarize(slp_grouped, sleep_mean = mean(sleep))
```

# Create boxplots of sleep across treatment groups
```{r}
ggplot(slp_reduced, aes(y = sleep, group = cond.f, color = cond.f)) +
  geom_boxplot() 
```



# Conduct an ANOVA
```{r}
lm1 <- lm(sleep ~ cond.f, data = slp_reduced)
anova(lm1)
```


## Obtain eta squared
```{r}
22892 / (22892 + 65385)
```



# Conduct planned contrasts
## Create contrasts
```{r}
c1 <- c(-1, 0, 1)
c2 <- c(0, -1, 1)
```



## Bind contrasts to predictor variable 
```{r}

contrasts(slp_reduced$cond.f) <-  cbind(c1,c2)

```

# View the structure of the data
```{r}
#View the structure and attributes
str(slp_reduced)

```


## View the attributes of slp_reduced$cond.f
```{r}
attributes(slp_reduced$cond.f)
```



# Run ANOVA with contrasts
```{r}
lm_contrast <- lm(sleep ~ cond.f, data = slp_reduced)
summary.lm(lm_contrast)
```


