---
title: "Module 6 Demo"
output: 
  html_notebook:
    toc: yes
---


# Load Libraries
```{r,message=FALSE}
library(ppcor)
library(psych)
library(tidyverse)
library(olsrr)
library(apaTables)
```

# read in data
```{r}
bac_a <- read_csv("bac_a.csv")
bac_b <- read_csv("bac_b.csv")

```

# Join (merge) our datasets
```{r}

bac_joined <- left_join(bac_a, bac_b, by = "id")

```


# select variables of interest
```{r}
bac_reduced <- select(bac_joined, weight, alc_gm, bac)
```


# Describe the data
```{r}
describe(bac_reduced)
```



# Make a scatterplot
```{r}
ggplot(bac_reduced, aes(x = alc_gm, y = bac)) +
  geom_point() +
  geom_smooth(method = "lm")
```


# Get correlation of variables
## Via cor() function
```{r}
cor(bac_reduced)
```

## apa.cor.table()
```{r}
apa.cor.table(bac_reduced, "bac_correlations.doc", show.conf.interval = TRUE)
```

## Get partial correlation
```{r}

pcor.test(x = bac_reduced$alc_gm, y = bac_reduced$bac, z = bac_reduced$weight)

```



## Get semipartial correlation
```{r}

spcor.test(x = bac_reduced$alc_gm, y = bac_reduced$bac, z = bac_reduced$weight)

```



# Fit a Simple Linear Regression (SLR) Model
```{r}
mod1 <- lm(bac ~ alc_gm, data = bac_reduced)
```


## Display SLR model results
### Via olsrr's ols_regress function
```{r}
ols_regress(mod1)
```
# Plot the regression
## Simple plot
```{r}
ggplot(bac_reduced, aes(x = alc_gm, y = bac)) +
  geom_point() +
  geom_smooth(method = "lm")
```


## Complex plot for intercept proof
```{r}
ggplot(bac_reduced, aes(x = alc_gm, y = bac)) +
  geom_point() +
  geom_smooth(method = "lm", fullrange = TRUE) +
  scale_x_continuous(limits = c(-1, 60)) +
  scale_y_continuous(limits = c(0, .15)) +
  geom_vline(xintercept = 0, color = "red") +
  geom_hline(yintercept = .003, color = "red") +
  annotate(geom = "text", x = 6.8, y = .003, label = "<- Intercept = .003")


```

