---
title: "Factorial ANOVA"
output: 
  html_notebook:
    toc: yes
 
  
---
# Load Libraries 

```{r}
library(psych)
library(tidyverse)
library (apaTables)
library(car)
install.packages("lsr")
library(lsr)
library(MBESS)
```

# Import data

```{r}
slp<-read.csv("slpdata.csv")
```
# Factor the grouping variables
```{r}
slp <- mutate(slp, 
                 sex.f = factor(sex,
                                   levels = c(2,1),
                                   labels = c("female", "male")))
slp <- mutate(slp, 
              cond.f = factor(cond,
                              levels = c(1,2,3),
                              labels = c("self-help", "group-based", "group + partner")))
```

# Calculate descriptives
## For whole dataset

```{r}
describe(slp)
```

## Descriptives by grouping variables
```{r}
slp_grouped <- group_by(slp, cond.f, sex.f)
summarize(slp_grouped, mean_sleep = mean(sleep, na.rm = TRUE))
```
# Visualize the data
## Create boxplots of sleep efficiency across treatment groups
```{r}
ggplot(slp, aes(y = sleep, color = cond.f)) + geom_boxplot()
```
## Create boxplots of sleep efficiency across sex
```{r}
ggplot(slp, aes(y=sleep, color = sex.f)) + geom_boxplot()
```
## Create boxplots of sleep efficiency across treatment groups and sex
```{r}
ggplot(slp, aes(y=sleep, color = cond.f)) + geom_boxplot() + facet_wrap(~sex.f)
```
# Conduct a factorial ANOVA

```{r}
model <-lm(sleep~ cond.f + sex.f + cond.f*sex.f, data = slp)
Anova(model, type = 3)
```
The F statistic for each predictor and the interaction is significant.
# Output the model in apaTables
```{r}
apa.aov.table(model, filename = "Table1.doc", conf.level = .95)
```
The partial eta squared is .14 for condition, .16 for sex, and .01 for the interaction of the two. Partial eta squared is an effect size.

# Calculate partial eta^2 confidence intervals via MBESS' ci.pvaf() function
## for sex.f
```{r}
ci.pvaf(F.value = 49.33, df.1 = 1, df.2 = 594, N = 600, conf.level = .95)
```
## for cond.f
```{r}
ci.pvaf(F.value = 115.437, df.1=2, df.2 = 594, N = 600, conf.level = .95)
```
## Cond.f*sex.f
```{r}
ci.pvaf(F.value = 3.7865, df.1=2, df.2 = 594, N = 600, conf.level = .95)
```
# Create an interaction plot
```{r}
slp_grouped<-group_by(slp, sex.f, cond.f)
slp_graph <-summarize(slp_grouped, sleep = mean(sleep))
slp_graph <-ungroup(slp_grouped)

ggplot(slp_graph, aes(x= sex.f, y = sleep, color = cond.f)) + geom_line(aes(group = cond.f)) + 
  scale_y_continuous(limits = c(25,100)) +
  labs(title = "Interaction plot", subtitle = "sleep by Sex and Condition")
```

# Set up contrasts
## Compare male (1) to female (-1)
```{r}
#sex
sc1<- c(1,-1)
contrasts(slp$sex.f)<-sc1

```

## Compare self-help(1) to group-based(-1)
```{r}
cc1<- c(1,-1, 0)
```
## Compare self-help(1) to group-based+partner(-1)
```{r}
cc2<- c(1, 0, -1)

contrasts(slp$cond.f)<-cbind(cc1,cc2)
```

# Run ANOVA with contrast statements
## summary.lm() to analyze constructs

```{r}
model_contrasts <- lm(sleep ~sex.f*cond.f, data = slp)
summary.lm(model_contrasts)
```
GM = 69.83 is different from 0.
There is a significant difference between males and females. There is no significant difference between self help and group based conditions. There is a significant difference between self help and group based + partner. There is no difference based on sex between self-help and group based groups.
There is a difference based on sex between self-help and group based + parter groups.
























