---
title: "R Notebook"
output: html_notebook
---



# Load libraries
```{r, message = FALSE}
library(tidyverse)
library(psych)
library(car)
library(lsr)
library(MBESS)
library(apaTables)
```






# Import Data
```{r, message = FALSE}

diet <- read_csv("Diet.csv")

```


# Factor grouping variables
```{r}
diet <- mutate(diet,
  gender.f = factor(gender, levels = c(0,1), labels = c("female", "male")))

diet <- mutate(diet, Diet.f = factor(Diet, levels = c(1,2,3), 
                                   labels = c("Control", "Atkins", "Nutrisystem")))
```


# Select variables of interest 
```{r}
diet_reduced <- select(diet, gender.f, Diet.f, weightLOST)
```


# Calculate descriptives
# For whole dataset
```{r}
describe(diet_reduced)
```


```{r}
diet_grouped <- group_by(diet_reduced,  gender.f, Diet.f)
summarize(diet_grouped, mean_weightLOST = mean(weightLOST))
```


# Visualize the data
## Create boxplots of weightloss across gender
```{r}
ggplot(diet_reduced, aes(y = weightLOST, color = gender.f)) +
  geom_boxplot() 

```



## Create boxplots of weightloss across diet types
```{r}
ggplot(diet_reduced, aes(y = weightLOST, color = Diet.f)) +
  geom_boxplot() 

```



## Create boxplots of weightloss across diet groups and gender
```{r}
ggplot(diet_reduced, aes(y = weightLOST, color = Diet.f)) +
  geom_boxplot() +
  facet_wrap(~gender.f)

```




# Conduct a factorial ANOVA
```{r}
model <- lm(weightLOST ~ gender.f*Diet.f, data = diet_reduced)
Anova(model, type = 3)
```


# View and output ANOVA table via apaTables's apa.aov.table()
```{r}
apa.aov.table(model, filename = "model.doc", conf.level = .95)
```


## Calculate 95% CIs
### gender.f
```{r}

ci.pvaf(F.value = 24.43, df.1 =1, df.2 = 72, N = 78, conf.level = .95)

```

### Diet.f
```{r}
ci.pvaf(F.value =  10.14, df.1 =2, df.2 = 72, N = 78, conf.level = .95)
```


### gender.f*Diet.f
```{r}
ci.pvaf(F.value =  3.79, df.1 =2, df.2 = 72, N = 78, conf.level = .95)
```


## Interaction plot
```{r}

diet_grouped <- group_by(diet_reduced, gender.f, Diet.f)
diet_graph <- summarize(diet_grouped, weightLOST = mean(weightLOST))
diet_graph <- ungroup(diet_graph)


ggplot(diet_graph, aes(x = gender.f, y = weightLOST, color = Diet.f)) +
  geom_line(aes(group = Diet.f)) + 
  scale_y_continuous(limits = c(0,6)) +
  labs(title = "Interaction plot", subtitle = "WeightLOST by Gender & Diet") 

  

```



## Set up contrasts
```{r}

#Contrasts for gender
gc1 <- c(-1,1)

#Contrasts for diet type
dc1 <- c(1,-1,0)
dc2 <- c(1,0,-1)


contrasts(diet_reduced$gender.f) <- gc1
contrasts(diet_reduced$Diet.f)   <- cbind(dc1, dc2)
```


### summary.lm()
```{r}

model_contrasts <- lm(weightLOST ~ gender.f*Diet.f, data = diet_reduced)
summary.lm(model_contrasts)
```

