---
title: "R Notebook"
output: html_notebook
---

# Load libraries
```{r}
library(psych)
library(tidyverse)
library(apaTables)
library(car)
library(lsr)
library(MBESS)
```

# Import data
```{r}
slp <- read_csv("slpdata.csv")
```

# Factor the grouping variable
```{r}
slp <- mutate(slp, 
              sex.f = factor(sex, 
                                  levels = c(1,2),
                                  labels = c("male", "female")),
              cond.f = factor(cond, 
                              levels = c(1,2,3),
                              labels = c("self-help", "group-based", "group + partner")))
```


# Visualize the data
## Create boxplots of sleep efficiency across treatment groups
```{r}
ggplot(slp, aes(y = sleep, color = cond.f)) +
  geom_boxplot()

```


## Create boxplots of sleep efficiency across sex
```{r}
ggplot(slp, aes(y = sleep, color = sex.f)) +
  geom_boxplot()
```

## Create boxplots of sleep efficiency across groups & sex
```{r}
ggplot(slp, aes(y = sleep, color = cond.f)) +
  geom_boxplot() +
  facet_wrap(~sex.f)
```
# Conduct Factorial ANOVA
```{r}

mod <- lm(sleep ~ sex.f*cond.f, data = slp)
Anova(mod, type = 3)


```

# Output an APA styled ANOVA via apaTables' apa.aov.table() function
```{r}
apa.aov.table(mod, filename = "model.doc", conf.level = .95)
```


```{r}
ci.pvaf(F.value = 115.44, df.1 = 1, df.2 = 594, N = 600, conf.level = .95)
ci.pvaf(F.value = 113.84, df.1 = 2, df.2 = 594, N = 600, conf.level = .95)
ci.pvaf(F.value =   3.79, df.1 = 2, df.2 = 594, N = 600, conf.level = .95)
```


# Run contrasts
```{r}
sc1 <- c(1,-1)

cc1 <- c(1, -1, 0)
cc2 <- c(1, 0, -1)

contrasts(slp$sex.f) <- sc1
contrasts(slp$cond.f) <- cbind(cc1, cc2)



attributes(slp$sex.f)
attributes(slp$cond.f)

```


# Run ANOVA with contrast statements
```{r}
mod_contrasts <- lm(sleep ~ sex.f*cond.f, data = slp)
summary.lm(mod_contrasts)
```

# Optional interaction plots

```{r}
slp_grouped <- group_by(slp, sex.f, cond.f)
slp_graph   <- summarize(slp_grouped, sleep = mean(sleep, na.rm = TRUE))
slp_graph   <- ungroup(slp_graph)


ggplot(slp_graph, aes(x = sex.f, y = sleep, color = cond.f)) +
  geom_line(aes(group = cond.f)) 
```



```{r}
interaction.plot(x.factor = slp$sex.f, trace.factor= slp$cond.f, response =slp$sleep, type="b")
```





