---
title: "Answers"
output: 
  html_notebook:
    toc: yes
---

# Load libraries
```{r}
library(tidyverse)
library(epitools)
```

# Read in data
```{r}
slp <- read_csv("slp_cat.csv")
```

# Get structure of the data
```{r}
str(slp)
```


# Get sumary of data
```{r}
summary(slp)
```

# Factor variables
```{r}
slp <- mutate(slp, 
              sex       = as.factor(sex),
              cond      = as.factor(cond),
              prior     = as.factor(prior),
              sleep = as.factor(sleep)
              )
```


# Get structure of data again
```{r}
str(slp)
```


# Get summary of data again
```{r}
summary(slp)
```


# Run a chi square of sex and sleep cat
## Step 1: Create a table of your variable
```{r}
sex_sleep <- select(slp, sex, sleep)
sex_sleep_table <- table(sex_sleep)

sex_sleep_table
```

## Step 2: run chisq.test() on the new table object
```{r}
chisq.test(sex_sleep_table)
```
### Step 2.1: If significant, calculate the contigency coeffient effect size
Results were significant (p <.05)
```{r}
# contingency coefficient = sqrt(X^2 / (X^2 + N))
sqrt(82.564 / (82.564 + 600))
```


# Get odds ratio of our 2x2 table (sex_sleep_table)
## First calculate it by hand
### OR of Males being unhappy as compared to females

```{r}
# Get our table
sex_sleep_table
```
```{r}

#Calulate Odds for Males being happy
males_bad <- 230 / 67

# Calculate odds for males being unhappy
males_good <- 123 / 180

# Get the odds ratio
OR <-  males_good / males_bad
OR

```
Males are .19 times as likely to have good sleep as compared to females (meaning they are likely to have a worst sleep)

### Then calculate it via oddsratio.wald()
```{r}

oddsratio.wald(sex_sleep_table)

```

Males are .19 times as likely to have good sleep as compared to females (meaning they are likely to have a worst sleep)

