---
title: "Analysis of categorical variables"
output: 
  html_notebook:
    toc: yes
---

# Load Libraries
```{r}
library(tidyverse)
library(epitools)
```

# Read in data
```{r}
gss <- read_csv("gss.csv")
```

# reduce dataset to variables of interest
```{r}
gss_reduced <- select(gss, id, sex, happy, zodiac, party)
```


# Get structure of dataset
Notice that you have 4 character (chr) variables. These are going to give us problems later if we don't do something about it. 
```{r}
str(gss_reduced)
```

# Get summary of character variables
Because most of our variables are character, the summary function is basically useless.
```{r}
#provides us useless output
summary(gss_reduced)
```

# Factor our character variables via mutate()
Factoring provides a numeric value to a character variable.
```{r}

gss_reduced <- mutate(gss_reduced, 
                      sex = as.factor(sex),
                      happy = as.factor(happy),
                      zodiac = as.factor(zodiac),
                      party = as.factor(party)
                      )

```


# Get new structure of dataset
Notice, our character variables are now factors
```{r}
str(gss_reduced)
```

# Get new summary of dataset
"maxsum =" indicates how many factor levels you want to show (the default deprecates at 6 levels)
```{r}
summary(gss_reduced, maxsum = 12)
```



# chi-square analyses:
## Step 1: Create a table of your variables
Below we will be creating a table object of the variables sex & happy from the gss_reduced dataset.
We will then view the object.
```{r}


sex_happy <- select(gss_reduced, sex, happy)
sex_happy_table <- table(sex_happy)

sex_happy_table
```
## Step 2: run chisq.test() on the new table object
```{r}
chisq.test(sex_happy_table)
```
### Step 2.1: If significant, calculate the contigency coeffient effect size
Results were significant (p <.05)
```{r}
# contingency coefficient = sqrt(X^2 / (X^2 + N))
sqrt(8.4605 / (8.4602 + 149))
```
# Another chi-square analysis 
## Step 1: Create a table of your variables
Below we will be creating a table object of the variables sex & party from the gss_reduced dataset.
We will then view the object.
```{r}

sex_party <- select(gss_reduced, sex, party)
sex_party_table <- table(sex_party)

sex_party_table
```
## Step 2: run chisq.test() on the new table object
```{r}

chisq.test(sex_party_table)

```
### Step 2.1: If significant, calculate the contigency coeffient effect size
Results were non-significant ( p > .05)


# Get odds ratio of our 2x2 table (sex_happy_table)
## First calculate it by hand
### OR of Males being unhappy as compared to females

```{r}
# Get our table
sex_happy_table
```

```{r}

#Calulate Odds for Males being happy
males_happy <- 37 / 78

# Calculate odds for males being unhappy
males_unhappy <- 21 / 13

# Get the odds ratio
OR <-  males_unhappy / males_happy
OR

```
Males are 3.405 times to be UNHAPPY as compared to females. 

## Then calculate it via oddsratio.wald()
The oddsratio.wald() function comes from the epitools package.
```{r}

oddsratio.wald(sex_happy_table)

```
Males are 3.405 times to be UNHAPPY as compared to females. 


### Reverse columns and get OR of men being HAPPY as compared to Females
```{r}
oddsratio.wald(sex_happy_table, rev = "columns")
```

Males are .29 times as likely to be happy as compared to females (meaning they are less likely to be happy!)





