---
title: "Category_variables_notebook"
output:
  html_document:
    df_print: paged
  html_notebook: 
    toc:yes
---
#Load libraries
```{r}
install.packages("tidyverse")
library(tidyverse)

install.packages("epitools")
library(epitools)
```
#Import data
```{r}
slp<- read.csv("slp_cat.csv")
```
#See structure of data frame
```{r}
str(slp)
```
#Get a summary of the dataset
```{r}
summary(slp)
```
My variables are in character format and I need
to convert them to factors

#factor the variables via mutate
```{r}
slp <- mutate(slp, 
              sex=as.factor(sex),
              cond=as.factor(cond),
              prior=as.factor(prior),
              sleep=as.factor(sleep))
```
#See the structure of the dataset again
```{r}
str(slp)
```
#Get a summary of the dataset again
```{r}
summary(slp)
```

#Run a chi-square of sex and sleep variables
##Step 1: Create a frequency table of your variable
```{r}
sex_sleep <- select(slp, sex, sleep)
```


```{r}
sleep_sex_table <- table(sex_sleep)

sleep_sex_table
```

## run chisq.test() on the new frequency table object
```{r}
chisq.test(sleep_sex_table)
```
There is a significant difference between groups as the P value is less that .05
##Step 2.1: Calculate the contingency coefficient effect size
```{r}
sqrt(82.564 / (82.564 + 600))
```
This means there is a moderate effect size; not a huge difference between groups

#Get odds ratio of our sex_sleep_table
##calcuate by hand

.199

#Calculate it via oddsratio.wald() from the epitools package
```{r}
oddsratio.wald(sleep_sex_table)
```

