---
title: "Wtloss_notebook"
author: "Jordan Thomas"
output:
  html_document:
    df_print: paged
---

#Load Libraries
```{r, warning=FALSE}
library(ppcor)
library(psych)
library(olsrr)
library(apaTables)
library(tidyverse)
```

#Import Data
```{r}
wtlossa <- read_csv("wtloss_parta.csv")
wtlossb = read_csv("wtloss_partb.csv")
```

#Merge Data
```{r}
wtloss = merge.data.frame(wtlossa, wtlossb)
```

#Describe Data
##Via base R's summary function
```{r}
summary(wtloss)
```

##Via describe function
```{r}
describe(wtloss)
```

#Check for outliers
```{r}
##Via Base R's boxplot function
boxplot(wtloss$lbslost)
##Via ggplot2's ggplot function
ggplot(wtloss, aes(y = lbslost)) + geom_boxplot()
```
Based on the boxplots above, it appears that there is one outlier.

#Plot relationships between variables
```{r}
##Via Base R's plot function
plot(wtloss[ , c("lbslost", "caldef", "selfeff")])
```

```{r}
##Via ggplot2's ggplot function
ggplot(wtloss, aes(x = lbslost, y = caldef)) + geom_point() +
geom_smooth(method = "lm")

ggplot(wtloss, aes(x = lbslost, y = selfeff)) + geom_point() +
geom_smooth(method = "lm")

ggplot(wtloss, aes(x = selfeff, y = caldef)) + geom_point() +
geom_smooth(method = "lm")
```

#Correlations
##Pearson's Correlation Matrices
```{r}
###Via base R's cor function
cor(wtloss)
```

```{r}
###Via apaTables' apa cor.table function
apa.cor.table(wtloss, "Weight Loss Table.doc", show.conf.interval = TRUE)
```

There is a strong positive correlation between the amount of pounds lost and calorie deficit (r = .72, p < .01). There is also a strong positive correlation between self-efficacy and and pounds lost (r = .54, p < .01). Additionally, there is a strong positive correlation between calorie deficit and self-efficacy (r = .59, p < .01).

#Get Partial Correlation
```{r}
attach(wtloss)
pcor.test(x = lbslost, y = caldef, z = selfeff)
```

The results of the partial correlation indicate that pounds lost has a strong, positive correlation to calorie deficit (r = .59, p < .001), even when controlling for self-efficacy.

#Get Semipartial Correlation
```{r}
spcor.test(lbslost, caldef, selfeff)
```

The results of the semi-partial correlation shows that pounds lost and calorie deficit have a strong, positive relation (r = .5, p < .001), even when controlling for self-efficacy.

#Fit a Simple Linear Regression (SLR) Model
```{r}
mod1 = lm(lbslost ~ caldef, data = wtloss)
```

##Display SLR Model Results
```{r}
###Via base R's summary function
summary(mod1)
```

```{r}
###Via olsrr's ols_regress function
ols_regress(mod1)
```

The y-intercept for the model is -.12. The unstandardized beta coefficient for caldef is .29. The standard error of caldef is .03. The results indicate there was a significant relation between caldef and lbslost (b = .29, p < .001). Overall, the caldef was able to account for a significant amount of variance in lbslost (R^2 = .52, p < .001).