---
title: "gapminder data wrangling and visualization"
output: 
  html_notebook:
    toc: yes
---


# Load libraries
```{r}
library(tidyverse)
```


# Read in data
```{r}

gm <- read_csv("gapminder.csv")

```

# filter() the dataset

There are many ways to filter data. To filter data, we use "logical operators" within the filter() function. 

a.	 `==` Which means “Is exactly equal to”
b.	`!=` Which means “Is NOT equal to”
c.	`>` Which means “Is greater than”
d.	`<` Which means “Is less than”
e.	`>=` Which means “Is greater than OR equal to”
f.	`<=` Which means “Is less than OR equal to”
g.	`is.na()` Which means “Contains an NA value”.
h.	`!is.na()` Which means “Does NOT contain an NA value”
i.	`%%` - is called a “modulo”, which means “Has a remainder of” (It’s not used as much, but is actually very useful)


```{r}

#Filter to only years greater than 1970
filter1 <- filter(gm, year > 1970)

#Filter to when year is 2007 & continent is Asia. A "&" means AND!
filter2 <- filter(gm, year == 2007 & continent == "Asia")

#Filter to only include data where the continent is "Americas" OR "Europe"
filter3 <- filter(gm, continent == "Americas" | continent == "Europe")


```


# select() the dataset

We use the select function to choose columns in our dataset.

```{r}

#Keep only country, year, lifExp
select1 <- select(gm, country, year, lifeExp)

# Remove just one of the variables (And keep the rest)
select2 <- select(gm, -lifeExp)

#select() is also great for just rearranging variables. putting "everything()" at the end of the function tells select you want to keep all of the variables. 
select3 <- select(gm, country, continent, gdpPercap, everything())

```


# mutate() the dataset

We use the mutate() function to create new variables or modify existing variables. 

We can multiply variables (*)  
We can divide variables (/)  
We can add variables (+)  
We can subtract variables (-)  
We can exponentiate variables (**)  

```{r}




#Multiply gdpPercap and population. Save it into a new variable called "gdp"
mutate1 <- mutate(gm, gdp = gdpPercap*pop)


#divide gdpPercap by 10. Save it into a new variable called gdp_10
mutate2 <- mutate(gm, gdp_10 = gdpPercap/10)



```



# arrange() the dataset

You can use arrange() to sort a dataframe

```{r}
#sort the dataframe by year
arrange1 <- arrange(gm, year)

#Sort two by two variables (year then lifeExp)
arrange2 <- arrange(gm, year, lifeExp)

#Sort in descending order
arrange3 <- arrange(gm, desc(year))

#combine it all
arrange4 <- arrange(gm, year, desc(lifeExp))

```


# summarize() the dataset

you can summarize the dataset in certain ways. 

## Get the mean of the lifeExp variable.
```{r}


summarize(gm, avg_lifeExp = mean(lifeExp, na.rm = TRUE))

```

## Get the mean of lifeExp, grouped by year
```{r}

#First group your variable by year
gm_by_year <- group_by(gm, year)

#Optional: Check the grouping structure of the dataframe
#group_vars(gm_by_year)

#Summarize the variable
summarize(gm_by_year, avg_lifeExp = mean(lifeExp, na.rm = TRUE), .groups = "drop")


#Summarize multiple aspects by year
gm_by_year <- group_by(gm, year)
summarize(gm_by_year, 
          avg_lifeExp = mean(lifeExp), 
          sd_lifeExp = sd(lifeExp), 
          ncount = n(), .groups = "drop")

#Ungroup the dataset 
gm_by_year <- ungroup(gm_by_year)
#group_vars(gm_by_year)



```

# Using the "Pipe" operator (%>%)
You can use pipes to combine any of these functions (i.e. filter(), select()) together and make your code more readable. 

# Let's filter() & select()
```{r}
indo <- gm %>%
  filter(country == "Indonesia") %>%
  select(year, lifeExp, country)
```


## A bit more complex... filter(), select(), group_by, & summarize()
```{r}
gm %>%
  filter(continent != "Oceania") %>%
  select(year, continent, country, lifeExp) %>%
  group_by(continent, year) %>%
  summarize(avg_lifeExp = mean(lifeExp), 
            sd_lifeExp = sd(lifeExp), 
            ncount = n(), 
            .groups = "drop")
```



# Visualizing data with ggplot()
## There are 3 essential components to ggplot

1.) The dataset  
2.) Your aesthetics function (aes())  
3.) your geom_ layer(s)  

# Create a histogram
```{r}
ggplot(gm, aes(x = lifeExp)) +
  geom_histogram(fill = "dark green")
```



## Add a title to my histogram via labs() layer
```{r}
ggplot(gm, aes(x = lifeExp)) + 
  geom_histogram(fill = "dark green") +
  labs(title = "This is my title", subtitle = "This is my subtitle")
```

# Create a scatterplot
## A simple one
```{r}
ggplot(gm, aes(x = gdpPercap, y = lifeExp)) +
  geom_point()
```

## A more complex one
```{r}
ggplot(gm, aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ continent) +
  scale_x_log10() +
  labs(title = "This is the title of my chart", 
       subtitle = "Here is my subtitle")
```



# Create a bar chart
```{r}
ggplot(gm, aes(x = continent, fill = continent)) + 
  geom_bar() + 
  labs(title = "How many data points do we have for each continent?", subtitle = "That's pretty neat!")
```

```{r}
ls(pattern = '^geom_', env = as.environment('package:ggplot2'))

ggplot2:::.all_aesthetics
```



