---
title: "Module 7 Lab Activity"
subtitle: "PSY 652 Research Methods"
date: "Oct 14, 2020"
output: 
  html_document:
    theme: readable
---

* 1.	Create a new R notebook from your project file and name it “slr_inference”

* 2.	Create a new R chunk with a first level header: “Load Libraries” 

  - a.	load the packages in this R chunk: tidyverse and psych

* 3.	Create a new R chunk with the first level header: “Simulate the population”

  - a.	In your R chunk Copy and paste the following code:

```{r, eval = FALSE}
# generate some data
set.seed(1870)
height  <- rnorm(n=50000, m=70, sd=3.00) # generate a population of 50,000 (Avg Height of Men in the United States in inches)
my_pop <- data.frame(height)
```

*          

  - b.	Run the chunk. You should now have a dataset with 50,000 observations of 1 variable available in your environment. This is your simulated population.

* 4.	Insert a new R chunk with the first level header: “Describe data”

  - a.	In the R chunk, use the describe() function to summarize your dataset object (meanpop)

* 5.	Create a first level header that says: “Create a histogram of the population” 

  - a.	Use ggplot to create a histogram of height. Set the histogram binwidth to 0.30. 
  - b.	Additionally, add the following:
    - i.	A title that says: “Distribution of height for U.S. Men"
    - ii.	A subtitle that says: “Simulation of a population”
    - iii.	Label the x-axis as: “Height” 

Hint: 
```{r, eval = FALSE, error = TRUE}

ggplot(data, aes(x = var)) + 
  geom_histogram(binwidth = .30) + 
  labs(title = "Title here", subtitle = "subtitle here", x = "x-axis here")


```


* 6.	Create a first level header: “Randomly select 3 samples of 100 from the population”

  - a.	In an R chunk use the sample_n() function to randomly pull 100 participants from your population dataset with replacement. Save these each into their own object called s1, s2, and s3. 
  
* 7.	Create a first level header: “Describe each of the 3 samples”
  - a.	In an R chunk, use the describe() function three times to describe each of the sample datasets you just made (s1, s2, & s3)
    - i.	Compare these to the original population summary. Do the numbers look relatively similar? Although not exact, the mean value should be close to the true population mean. 
* 8.	Create a first level header: Inference for SLR 
* 9.	Create a second level header: “Simulate the population SLR results
  - a.	In an R chunk, copy and paste the following code:
  
```{r, eval = FALSE}
# Generate some data 
set.seed(83587) 
weight <- rnorm (mean = 197.0, sd = 1.2, n = 50000) #Create weight var Average weight of men in the U.S. in pounds

b0 <- 0 # Pop intercept
b1 <- 1 # Pop slope
sigma <- 1.0 # Pop SD

e <- rnorm(mean = 0, sd = sigma, n = 50000) #generate error

height <- b0 + b1*weight + e #create height variable

my_regpop <- data.frame(weight, height) #Dataset

```
    
*        
  - a.	Run the chunk. You should now have a dataset with 50,000 observations of 2 variable available in your environment. This is your simulated population for height and weight.

* 10.	Create a first level header: “Get a scatterplot of the population”
  - a.	In an R chunk, use ggplot to create a scatteplot with weight on the x-axis and height on the y-axis. Add a regression line (geom_smooth()) to the plot. Give the plot a snazzy title and label your x and y axes. 
  
Hint: 
```{r, eval = FALSE, error=TRUE}

ggplot(data, aes(x = var1, y = var2)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Snazzy title", x = "name1", y = "name2")  

```

* 11.	Create a second level header: “Fit a SLR model to the population”
  - a.	In an R chunk, regress height on weight. (Weight is your predictor, height is the outcome variable). Save it to a new object called “popmod”.
  - b.	In the same R chunk, display a summary of the model using summary()
  
* 12.	Create a second level header: “Randomly select 3 samples of 100 from the population”
  - a.	In an R chunk use the sample_n() function to randomly pull 100 participants from your population dataset with replacement. Save these each into their own object called ss1, ss2, and ss3. 

* 13.	Create a second level header called: “Run a SLR on each sample”
  - a.	Run a regression analysis on each newly created object (ss1, ss2, & ss3) with height regressed on weight (The same equation you did for your population dataset). Save them into their own unique objects called r1, r2, r3.     
  - b.	In the same R chunk, run summary() on each object (r1, r2, & r3). You will have 3 summary() statement calls. 
    - i.	View the output from each of the regression outputs. Are they similar? They are expected to be slightly different but will be around the population parameters.
    
* 14.	Once you’ve completed all of these steps, Restart R and Run All Chunks, and then preview your notebook. Save your notebook as both a .Rmd and an html file and exit RStudio.

* 15.	Upload both the .Rmd and html version of your notebook to the assignment called “Module 7 Lab Activity” on the course Canvas page.









